---
title: "Appendix B - Data Sources"
author: "Ethan Shen, Malavi Ravindran, Steven Herrera Tenorio, Anna Darwish"
geometry: "left=1.25cm,right=1.25cm,top=1.3cm,bottom=1.3cm"
fontsize: 12pt
output: 
  pdf_document:
     number_sections: true
---

```{r}
library(pander)
```

# 2012 2020 Presidential Election & Electoral College Vote (1.1 and 1.2)

The Presidential Election & Electoral College Vote model (Model 1 from Appendix A) used the following dataset: `2020 US presidential election polls - all_polls.csv`.

```{r}
head(read_csv("2020 US presidential election polls - all_polls.csv"))
```

The following variables were used in Model 1 from Appendix A: states <- president_polls$state %>% unique
y <- president_polls$pct
r <- match(president_polls$state,states)
t <- president_polls$days_to_election + 1 
N_polls <- y %>% length
N_states <- states %>% length
N_days <- t %>% max
I_states <- diag(N_states)

table of datasources, aech row is a datasource,  the variables you get from it, and a mapping of the models 
i.e. presidential poll  csv,  x y z variables, Model  1
hilary's election results, x  y z  variables,  Model 1

# US Senate Election & NC Senate Election (2)


# Who Votes (Interim Report) (3.1)

Our "Who Votes" model utilizes the following datasets. These datasets were both located on the cluster, and originally come from https://dl.ncsbe.gov/?prefix=data/. Note that they were joined by a covariate, voter_reg_num, that appeared in both and represented an individual's unique registration number. 

```{r}
Data <- c("NC Voter Registration Data", "NC Voter History Data")
Variables <- c("County, Congressional District, Race Code, Ethnic Code, Age, Party, Gender", "voted_party_cd")
Notes <- c("Age was rebinned into age_bin, and 'Race Code' and 'Ethnic Code' were combined into a singular race_ethnicity variable", "This variable was not directly used in the model, but was used to determine if an individual actually voted. Individuals who were registered but did not vote would not have information regarding the party that they voted for in 2016")
pander(data.frame(Data, Variables, Notes), caption = "Data: Who Votes (Interim Report)")
```


# Congressional Elections Prediction (3.2)

Our Congressional Election models utilize the same two datasets. Again, these datasets were both located on the cluster, and originally come from https://dl.ncsbe.gov/?prefix=data/. Note that they were joined by a covariate, voter_reg_num, that appeared in both and represented an individual's unique registration number. 

```{r}
Data <- c("NC Voter Registration Data", "NC Voter History Data")
Variables <- c("County, Congressional District, Race Code, Ethnic Code, Age, Party, Gender", "Voted Party Code")
Notes <- c("Age was rebinned into age_bin, and 'Race Code' and 'Ethnic Code' were combined into a singular race_ethnicity variable", "We use this as our outcome variable, as we are interested in predicting whether or not voters in NC will vote for the Democratic or Republican candidates")
pander(data.frame(Data, Variables , Notes), caption = "Data: Congressional Elections Prediction")
```

* NC Voter Registration Data: County, Congressional District, Race Code, Ethnic Code, Age, Party Affiliation (Note that in our model, "Age" was rebinned into age_bin, and "Race Code" and "Ethnic Code" were combined into a singular race_ethnicity variable)

* Note that they were joined by a covariate, voter_reg_num, that appeared in both and represented an individual's unique registration number. 



```{r}
model_num = c(
  rep("Model 1 - 2020 Presidential", 8),
  rep("Model 2 - 2012 Presidential", 8),
  rep("Model 3 - 2020 U.S. Senate", 8)
)
datasources = c(
  rep("2020 Presidential Polls (from 538)", 8),
  rep("2012 Presidential Polls (from The Economist)", 8),
  rep("2020 Senate Polls (from 538)", 8)
)

data_variables =  linebreak(c("states", "y", "r", "t", "N_polls", "N_states", "N_days", "I_states",
                              "states", "y", "r", "t", "N_polls", "N_states", "N_days", "I_states",
                              "states", "y", "r", "t", "N_polls", "N_states", "N_days", "I_states"))

descriptions = linebreak(c("Names of 15 Swing States (Each state has an ID 1 - 15)", "Biden's Vote Share from each Poll k (N=516)", "Matches State of Poll k to the State's Unique ID (N=516)", "Days to Election + 1 for Each Poll k (N=516)", "Number of Polls (N=516)", "Number of Swing States (N=15)", "Maximum Number of Days to Election + 1 (154 days)", "15 by 15 Identity Matrix",
                           
                           
                           "Names of 9 Swing States (Each state has an ID 1 - 9)", "Obama's Vote Share from each Poll k (N=400)", "Matches State of Poll k to the State's Unique ID (N=400)", "Days to Election + 1 for Each Poll k (N=400)", "Number of Polls (N=400)", "Number of Swing States (N=9)", "Maximum Number of Days to Election + 1 (95 days)", "9 by 9 Identity Matrix",
                           
                           "Names of 18 Swing Races (Each race has an ID 1 - 18)", "Democratic Candidate's Vote Share from each Poll k (N=506)", "Matches race of Poll k to the race's Unique ID (N=506)", "Days to Election + 1 for Each Poll k (N=506)", "Number of Polls (N=506)", "Number of Swing Races (N=18)", "Maximum Number of Days to Election + 1 (101 days)", "18 by 18 Identity Matrix"))

tibble(`Model Number` = model_num, 
       `Data Sources`=datasources,
       `Variables`=data_variables,
       Description=descriptions) %>% 
  kable(escape=F,   booktabs = T, align = "c",
        caption = "Data Sources",) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  collapse_rows(columns = c(1,2), latex_hline = "major") 
```

