---
title: "Appendix B - Data Sources"
author: "Ethan Shen, Malavi Ravindran, Steven Herrera Tenorio, Anna Darwish"
geometry: "left=1.25cm,right=1.25cm,top=1.3cm,bottom=1.3cm"
fontsize: 12pt
output: 
  pdf_document:
     number_sections: true
---

# Presidential Election & Electoral College Vote

The Presidential Election & Electoral College Vote model (Model 1 from Appendix A) used the following dataset: `2020 US presidential election polls - all_polls.csv`.

```{r}
head(read_csv("2020 US presidential election polls - all_polls.csv"))
```

The following variables were used in Model 1 from Appendix A: states <- president_polls$state %>% unique
y <- president_polls$pct
r <- match(president_polls$state,states)
t <- president_polls$days_to_election + 1 
N_polls <- y %>% length
N_states <- states %>% length
N_days <- t %>% max
I_states <- diag(N_states)

table of datasources, aech row is a datasource,  the variables you get from it, and a mapping of the models 
i.e. presidential poll  csv,  x y z variables, Model  1
hilary's election results, x  y z  variables,  Model 1

# US Senate Election & NC Senate Election


# NC Congressional Election

```{r}
model_num = c(
  rep("Model 1 - 2020 Presidential", 8),
  rep("Model 2 - 2012 Presidential", 8),
  rep("Model 3 - 2020 U.S. Senate", 8)
)
datasources = c(
  rep("2020 Presidential Polls (from 538)", 8),
  rep("2012 Presidential Polls (from The Economist)", 8),
  rep("2020 Senate Polls (from 538)", 8)
)

data_variables =  linebreak(c("states", "y", "r", "t", "N_polls", "N_states", "N_days", "I_states",
                              "states", "y", "r", "t", "N_polls", "N_states", "N_days", "I_states",
                              "states", "y", "r", "t", "N_polls", "N_states", "N_days", "I_states"))

descriptions = linebreak(c("Names of 15 Swing States (Each state has an ID 1 - 15)", "Biden's Vote Share from each Poll k (N=516)", "Matches State of Poll k to the State's Unique ID (N=516)", "Days to Election + 1 for Each Poll k (N=516)", "Number of Polls (N=516)", "Number of Swing States (N=15)", "Maximum Number of Days to Election + 1 (154 days)", "15 by 15 Identity Matrix",
                           
                           
                           "Names of 9 Swing States (Each state has an ID 1 - 9)", "Obama's Vote Share from each Poll k (N=400)", "Matches State of Poll k to the State's Unique ID (N=400)", "Days to Election + 1 for Each Poll k (N=400)", "Number of Polls (N=400)", "Number of Swing States (N=9)", "Maximum Number of Days to Election + 1 (95 days)", "9 by 9 Identity Matrix",
                           
                           "Names of 18 Swing Races (Each race has an ID 1 - 18)", "Democratic Candidate's Vote Share from each Poll k (N=506)", "Matches race of Poll k to the race's Unique ID (N=506)", "Days to Election + 1 for Each Poll k (N=506)", "Number of Polls (N=506)", "Number of Swing Races (N=18)", "Maximum Number of Days to Election + 1 (101 days)", "18 by 18 Identity Matrix"))

tibble(`Model Number` = model_num, 
       `Data Sources`=datasources,
       `Variables`=data_variables,
       Description=descriptions) %>% 
  kable(escape=F,   booktabs = T, align = "c",
        caption = "Data Sources",) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  collapse_rows(columns = c(1,2), latex_hline = "major") 
```

