---
title: "Senate Races"
output: html_document
---

Overview:
- 33 seats up; 12 dem, 23 republican
- currently there are 47 dems (counting independents) and 53 republicans 
- this means that, other than the 33 seats up, there are 35 democrats and 30 republicans in the senate
- to flip, democrats need to get up to 51. They need to win their 12 seats plus 4 of the republican seats! They have 8 guaranteed (not swing)
- DEMS NEED TO WIN 8 SWING STATES!


Seats up:

Georgia
Georgia*
Iowa
North Carolina
Montana
Maine
Kansa
South Carolina
Arkansas
Arizona*
Alabama
Michigan
Colorado
Texas
Mississippi
Minnesota
Kentucky
New Mexico
Louisiana
New Hampshire
Virginia
Oklahoma
Tennessee
Idaho
West Virginia
Oregon
New Jersey
South Dakota
Illinois
New England
Deleware
Maryland
Wyoming
Rhode Island
Arkansas

GUARANTEED: 8 DEM, 11 REP. Swing states? 16 total

BLUE
Alabama
Michigan
Minnesota
New Hampshire


RED:
Arizona*
Colorado
Georgia
Georgia*
Iowa
Kansas
Kentucky
Maine
Montana
North Carolina
South Carolina
Texas


```{r, function checking for installed packages, include=FALSE}
# Validate that all necessary packaged have been downloaded, install otherwise or throw err package DNE
pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,repos = "http://cran.r-project.org", dep=TRUE)
    if(!require(x,character.only = TRUE)) stop("Package not found")
  }
}
```

```{r package test, include=FALSE}
# Installing packages 
# Additionally, if you are struggling to download RJags, please visit the following link for help. It can also help to try to download the package locally or in RStudio Cloud: https://sites.google.com/a/utexas.edu/edm-principalstratification/downloading-installing-r-jags-rstudio
pkgTest("ggplot2")
pkgTest("R2jags")
pkgTest("tidyverse")
pkgTest("lubridate")

```

```{r load packages, include=FALSE, results="hide"}
library(ggplot2)
library(R2jags)
library(tidyverse)
library(lubridate)
```

```{r Read in Data}
#Filter to see if any non democratic or republican candidates are polling high

`%notin%` <- Negate(`%in%`)
senate_polls <- read_csv("senate_polls.csv")
senate_polls %>%
  filter(cycle == "2020") %>%
  filter(candidate_party %notin% c("DEM", "REP")) %>%
  arrange(desc(pct)) %>%
  select(poll_id, state, candidate_name, candidate_party, pct)

#TOP THREE: DUNBAR, HARRINGTON, WILSON
#Dunbar (AK) is a democrat, but running as an independent against republican and democratic candidates- but as of more recent polls he isn't relevant
#Harrington (AR) is liberatrian running against a republican
#Wilson (IL) is an independent running against a democrat- he can be counted as a democrat 

#filter out AR because there is a 0% chance a democrat wins here
#filter out IL because there is a 100% chance a democrat wins here
#remember to add back in one senator for IL because that's sort of a given!
```

```{r}
senate_polls_cleaned <- senate_polls %>%
  filter(cycle == "2020") %>%
  filter(population %in% c("lv","rv")) %>% 
  filter(office_type == "U.S. Senate") %>%
  
  #add separate elections for georgia and arizona special
  mutate(state = case_when(
    state == "Georgia" & seat_name == "Class III" ~ "Georgia*",
    state == "Arizona" & seat_name == "Class III" ~ "Arizona*",
    TRUE ~ as.character(state)
  )) %>%
  
  #get rid of arkansas because there is no chance that a democrat will win here 
  #get rid of illionis because basically guarantted a democrat wins
  filter(state %notin% c("Arkansas", "Illinois")) %>%
  filter(candidate_party %in% c("DEM", "REP")) %>%
  filter(poll_id != 58956 & question_id != 100828) %>%
  filter(poll_id != 67060 & question_id != 125304) #get rid of weird polls

```


```{r}
democratic_candidates <- senate_polls_cleaned %>%
  filter(candidate_party == "DEM") %>%
  group_by(question_id, poll_id, state, race_id, pollster, start_date, end_date) %>%
  summarize(total_dems_pct = sum(pct)) 
democratic_candidates 
```

```{r}
republican_candidates <- senate_polls_cleaned %>%
  filter(candidate_party == "REP") %>%
  group_by(question_id, poll_id, state, race_id, pollster, start_date, end_date) %>%
  summarize(total_rep_pct = sum(pct)) 
republican_candidates
```

```{r}
senate_data <- merge(democratic_candidates, republican_candidates, by = c("state", "question_id", "poll_id", "race_id", "pollster", "start_date", "end_date"))
senate_data <- senate_data %>%
  mutate(pct_dem = 100*total_dems_pct/(total_dems_pct + total_rep_pct))
senate_data
```

```{r}
senate_data
```

```{r}
senate <- senate_data %>%
   mutate(start_date = mdy(start_date),
         end_date = mdy(end_date),
         election_date = "2020-11-03" %>% as.Date(),
         days_to_election = election_date - end_date,
         y = pct_dem) %>%
  dplyr::filter(days_to_election <= 100) %>%
  filter(state %in% c("Alabama", "Arizona*", "Colorado", "Georgia*", "Georgia", "Iowa", "Kansas", "Kentucky", "Maine", "Michigan", "Minnesota", "Montana", "New Hampshire", "North Carolina", "South Carolina", "Texas"))
```


Model without priors from last senate election.

```{r}
states <- senate$state %>% unique
y <- senate$y
r <- match(senate$state,states)
t <- senate$days_to_election + 1 
N_polls <- y %>% length
N_states <- states %>% length
N_days <- t %>% max
jags_data <- list(y=y,t=t,r=r,
                  N_polls=N_polls,N_states=N_states,N_days=N_days)
```

```{r}
model_mv_senate <- function(){
  for(k in 1:N_polls){
    y[k] ~ dnorm(p[k],1/sigma2_y[r[k]]) #note no longer binomial
    p[k] = theta[r[k],t[k]]
  }
  for(j in 2:N_days){
    theta[1:N_states,j] ~ dmnorm(theta[1:N_states,j-1],Phi)
  }
  Phi ~ dwish(I_states,N_states+1) 
  Sigma = inverse(Phi)
  
  
  for(j in 1:N_states){
      sigma2_y[j] = 1/sigma2_y_inv[j]
      sigma2_y_inv[j] ~ dgamma(nu_y,nu_y*tau_y) 
      
      theta[j,1] ~ dnorm(mu0,pow(sigma2_0,-1))
  }
  nu_y ~ dunif(0,100)
  tau_y ~ dunif(0,100)
  
  nu_beta ~ dunif(0,100)
  tau_beta ~ dunif(0,100)
  
  mu0 ~ dnorm(50,pow(7.5,-2))
  sigma2_0 = 1/sigma2_0_inv
  sigma2_0_inv ~ dgamma(.5,.5)
}
```

```{r}
jags_data$I_states <- diag(N_states)

jags_sims_mv_senate <- jags(data = jags_data,model.file = model_mv_senate,parameters.to.save = c("theta","Sigma",
                                                                                "p","sigma2_y"),
                  n.iter = 5000)

```


```{r}
elec_sims <- jags_sims_mv_senate$BUGSoutput$sims.list$theta[1,,1]
vote_share_by_state <- data.frame(states, elec_sims)
```


```{r}
# this function finds the probability that the senate flips!
n_iter <- dim(jags_sims_mv_senate$BUGSoutput$sims.list$theta[,,1])


prediction_per_simulation_senate <- function(i){
  elec_sims <- jags_sims_mv_senate$BUGSoutput$sims.list$theta[i,,1]
  vote_share_by_state <- data.frame(states, elec_sims)

  dem_states_df <- vote_share_by_state %>%
    filter(elec_sims > 50)
  dem_states <- dem_states_df$states
  num_dem_states <- length(dem_states)
  return(num_dem_states)
  
}

final_results <- c()

for (i in 1:n_iter){
  final_results[i] = prediction_per_simulation_senate(i)
}

mean(final_results > 8)
mean(final_results)
```


Model WITH priors from last presidential election


```{r}
states <- senate$state %>% unique
y <- senate$y
r <- match(senate$state,states)
t <- senate$days_to_election + 1 
N_polls <- y %>% length
N_states <- states %>% length
N_days <- t %>% max
I_states <- diag(N_states)
```


```{r}
# swing states 
vote_share_2016 <- data.frame(senate$state, number = r)
vote_share_2016 <- unique(vote_share_2016) %>%
  mutate(clinton_percent = case_when(
    senate.state == "North Carolina" ~ 47/(47 + 51),
    senate.state == "Michigan" ~ 47/(47 + 48),
    senate.state == "New Hampshire" ~ 48/(48 + 47) ,
    senate.state == "Texas" ~ 43/(43 + 53) , 
    senate.state == "Minnesota" ~ 47/(47 + 45) ,
    senate.state %in% c("Georgia", "Georgia*") ~ 46/(46 + 51),
    senate.state == "Iowa" ~ 42/(42 + 52),
    senate.state == "Maine" ~ 48/(48 + 45),
    senate.state == "Arizona*" ~ 45/(48 + 45),
    senate.state == "Alabama" ~ 35/(35 + 63),
    senate.state == "Colorado" ~ 47/(47 + 44),
    senate.state == "Kansas" ~ 35/(35 + 63),
    senate.state == "South Carolina" ~ 41/(41 + 55),
    senate.state == "Kentucky" ~ 33/(33 + 63),
    senate.state == "Alabama" ~ 35/(35 + 63),
    senate.state == "Montana" ~ 36/(36 + 57),
    
    

  )) %>%
  mutate(clinton_percent = 100*clinton_percent)
h <- vote_share_2016$clinton_percent

```

```{r}
model_mv_senate_adj_priors <- function(){
  for(k in 1:N_polls){
    y[k] ~ dnorm(p[k],1/sigma2_y[r[k]]) #note no longer binomial
    p[k] = theta[r[k],t[k]] 
  }
  for(j in 2:N_days){
    theta[1:N_states,j] ~ dmnorm(theta[1:N_states,j-1],Phi)
  }
  
  Phi ~ dwish(I_states,N_states+1) 
  Sigma = inverse(Phi)
  #hierarchical component
  for(j in 1:N_states){
    sigma2_y[j] = 1/sigma2_y_inv[j]
    sigma2_y_inv[j] ~ dgamma(nu_y,nu_y*tau_y) 
      
    theta[j,1] ~ dnorm(h[j], pow(sigma2_0,-1))
    #theta[j,1] ~ dnorm(mu0,pow(sigma2_0,-1))
  }
  nu_y ~ dunif(0,100)
  tau_y ~ dunif(0,100)
  
    sigma2_0 = 1/sigma2_0_inv
  sigma2_0_inv ~ dgamma(2, 0.5)
}
```

```{r}
jags_data_senate_adj_prior <- list("y","t", "r", "N_polls", "N_states", "N_days", "h", "I_states")

jags_sims_mv_senate_adj_prior <- jags(data = jags_data_senate_adj_prior,model.file = model_mv_senate_adj_priors,parameters.to.save = c("theta","Sigma",
                                                                                "p","sigma2_y"),
                  n.iter = 5000)

```


```{r}
elec_sims <- jags_sims_mv_senate_adj_prior$BUGSoutput$sims.list$theta[1,,1]
vote_share_by_state <- data.frame(states, elec_sims)
```


```{r}
# this function finds the probability that the senate flips!
n_iter <- dim(jags_sims_mv_senate_adj_prior$BUGSoutput$sims.list$theta[,,1])


prediction_per_simulation_senate <- function(i){

  elec_sims <- jags_sims_mv_senate_adj_prior$BUGSoutput$sims.list$theta[i,,1]

  vote_share_by_state <- data.frame(states, elec_sims)
 
  dem_states_df <- vote_share_by_state %>%
    filter(elec_sims > 50)
  dem_states <- dem_states_df$states
  num_dem_states <- length(dem_states)
  return(num_dem_states)
  
}

final_results <- c()

for (i in 1:n_iter){
  final_results[i] = prediction_per_simulation_senate(i)
}

mean(final_results > 8)
mean(final_results)
```


