---
title: "Senate Races"
output: html_document
---

Overview:
- 35 seats up; 12 dem, 23 republican
- currently there are 47 dems (counting independents) and 53 republicans 
- this means that, other than the 35 seats up, there are 35 democrats and 30 republicans in the senate
- to flip, democrats need to get up to 51. They need to win their 12 seats plus 4 of the republican seats! They have 7 (VA, OR, IL,  NJ, DE,  CT, RI) guaranteed (not swing)
- DEMS NEED TO WIN 9 SWING STATES!


Seats up:

Georgia
Georgia*
Iowa
North Carolina
Montana
Maine
Kansa
South Carolina
Arkansas
Arizona*
Alabama
Michigan
Colorado
Texas
Mississippi
Minnesota
Kentucky
New Mexico
Louisiana
New Hampshire
Virginia
Oklahoma
Tennessee
Idaho
West Virginia
Oregon
New Jersey
South Dakota
Illinois
New England
Deleware
Maryland
Wyoming
Rhode Island
Arkansas

GUARANTEED: 8 DEM, 11 REP. Swing states? 16 total

8 Dem: Oregon, New Mexico, Illinois, VA, 

Delaware, Illinois, Mass, NJ NM, RI VA

BLUE
Alabama
Michigan
Minnesota
New Hampshire


RED:
Arizona*
Colorado
Georgia
Georgia*
Iowa
Kansas
Kentucky
Maine
Montana
North Carolina
South Carolina
Texas


```{r, function checking for installed packages, include=FALSE}
# Validate that all necessary packaged have been downloaded, install otherwise or throw err package DNE
pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,repos = "http://cran.r-project.org", dep=TRUE)
    if(!require(x,character.only = TRUE)) stop("Package not found")
  }
}
```

```{r package test, include=FALSE}
# Installing packages 
# Additionally, if you are struggling to download RJags, please visit the following link for help. It can also help to try to download the package locally or in RStudio Cloud: https://sites.google.com/a/utexas.edu/edm-principalstratification/downloading-installing-r-jags-rstudio
pkgTest("ggplot2")
pkgTest("R2jags")
pkgTest("tidyverse")
pkgTest("lubridate")
pkgTest("matrixStats")


```

```{r load packages, include=FALSE, results="hide"}
library(ggplot2)
library(R2jags)
library(tidyverse)
library(lubridate)
library(knitr)
library(matrixStats)

```

```{r Read in Data}
#Filter to see if any non democratic or republican candidates are polling high
# add New Mexico  and Missisipi 
`%notin%` <- Negate(`%in%`)
senate_polls <- read_csv("senate_polls.csv")
senate_polls$state %>% unique
senate_polls %>%
  filter(cycle == "2020") %>%
  filter(candidate_party %notin% c("DEM", "REP")) %>%
  arrange(desc(pct)) %>%
  select(poll_id, state, candidate_name, candidate_party, pct)

#TOP THREE: DUNBAR, HARRINGTON, WILSON
#Dunbar (AK) is a democrat, but running as an independent against republican and democratic candidates- but as of more recent polls he isn't relevant
#Harrington (AR) is liberatrian running against a republican
#Wilson (IL) is an independent running against a democrat- he can be counted as a democrat 

#filter out AR because there is a 0% chance a democrat wins here
#filter out IL because there is a 100% chance a democrat wins here
#remember to add back in one senator for IL because that's sort of a given!
```

```{r}
senate_polls_cleaned <- senate_polls %>%
  filter(cycle == "2020") %>%
  filter(population %in% c("lv","rv")) %>% 
  filter(office_type == "U.S. Senate") %>%
  
  #add separate elections for georgia and arizona special
  mutate(state = case_when(
    state == "Georgia" & seat_name == "Class III" ~ "Georgia*",
    state == "Arizona" & seat_name == "Class III" ~ "Arizona*",
    TRUE ~ as.character(state)
  )) %>%
  
  #get rid of arkansas because there is no chance that a democrat will win here 
  #get rid of illionis because basically guarantted a democrat wins
  filter(state %notin% c("Arkansas", "Illinois")) %>%
  filter(candidate_party %in% c("DEM", "REP")) %>%
  filter(poll_id != 58956 & question_id != 100828) %>%
  filter(poll_id != 67060 & question_id != 125304) #get rid of weird polls
senate_polls_cleaned$state %>% unique
```


```{r}
democratic_candidates <- senate_polls_cleaned %>%
  dplyr::filter(candidate_party == "DEM") %>%
  dplyr::group_by(question_id, poll_id, state, race_id, pollster, start_date, end_date) %>%
  dplyr::summarize(total_dems_pct = sum(pct)) 
democratic_candidates 
```

```{r}
republican_candidates <- senate_polls_cleaned %>%
  dplyr::filter(candidate_party == "REP") %>%
  dplyr::group_by(question_id, poll_id, state, race_id, pollster, start_date, end_date) %>%
  dplyr::summarize(total_rep_pct = sum(pct)) 
republican_candidates
```

```{r}
senate_data <- merge(democratic_candidates, republican_candidates, by = c("state", "question_id", "poll_id", "race_id", "pollster", "start_date", "end_date"))
senate_data <- senate_data %>%
  mutate(pct_dem = 100*total_dems_pct/(total_dems_pct + total_rep_pct))
senate_data
```

```{r}
senate_data$state %>% unique
```

```{r}
senate <- senate_data %>%
  mutate(start_date = mdy(start_date),
         end_date = mdy(end_date),
         election_date = "2020-11-03" %>% as.Date(),
         days_to_election = election_date - end_date,
         y = pct_dem) %>%
  #dplyr::filter(days_to_election <= 100) %>%
  filter(state %in% c("Alaska", "Alabama", "Arizona*", "Colorado", "Georgia*", "Georgia", "Iowa", "Kansas", "Kentucky",
                      "Maine", "Michigan", "Minnesota", "Mississippi",
                      "Montana", "New Hampshire",
                      "New Mexico",
                      "North Carolina", "South Carolina", "Texas")) #THIS is 538 (Kentucky+NH are  kinda red and kinda blue)

# filter(state %in% c("Alabama", "Arizona*", "Colorado", "Georgia*", "Georgia", "Iowa", "Kansas", "Kentucky",
#                     "Maine", "Michigan", "Minnesota", 
#                     "Montana", "New Hampshire",
#                     
#                     "North Carolina", "South Carolina", "Texas")) #  THIS is ballotpedia states
```


Model without priors from last senate election.

```{r}
states_senate <- senate$state %>% unique
y_senate <- senate$y
r_senate <- match(senate$state,states_senate)
t_senate <- senate$days_to_election + 1 
N_polls_senate <- y_senate %>% length
N_states_senate <- states_senate %>% length
N_days_senate <- t_senate %>% max
I_states_senate <- diag(N_states_senate)
jags_data_senate <- list(y=y_senate,t=t_senate,r=r_senate,
                         N_polls=N_polls_senate,N_states=N_states_senate,N_days=N_days_senate, I_states=I_states_senate)
```

```{r}
generate_senate_model_text = function(hierarchical_priors_senate) {
  model_file <- tempfile()
  likelihood_base_priors = "
  for(k in 1:N_polls){
    y[k] ~ dnorm(p[k],1/sigma2_y[r[k]]) #note no longer binomial
    p[k] = theta[r[k],t[k]]
  }
  for(j in 2:N_days){
    theta[1:N_states,j] ~ dmnorm(theta[1:N_states,j-1],Sigma)
  }
  
  Phi ~ dwish(I_states,N_states+1) 
  Sigma = inverse(Phi)
  
  
  for(j in 1:N_states){
      sigma2_y[j] = 1/sigma2_y_inv[j]
      sigma2_y_inv[j] ~ dgamma(nu_y,nu_y*tau_y) 
      
      theta[j,1] ~ dnorm(mu0, pow(sigma2_0),-1)
  }
  "
  
  model_text <- paste("model{ ", likelihood_base_priors, hierarchical_priors_senate,"}")
  writeLines(model_text,con=model_file)
  return(model_file)
}
```
ETHAN
sigma2_0_inv ~ dgamma(2,2), mu0 ~ dnorm(50, pow(7.5,-1))
sigma2_0_inv ~ dgamma(2,2), mu0 ~ dnorm(50, pow(10,-1))

STEVEN
sigma2_0_inv ~ dgamma(2,0.5), mu0 ~ dnorm(50, pow(7.5,-1))
sigma2_0_inv ~ dgamma(2,0.5), mu0 ~ dnorm(50, pow(10,-1))

ANNA
sigma2_0_inv ~ dgamma(5,5), mu0 ~ dnorm(50, pow(7.5,-1))
sigma2_0_inv ~ dgamma(5,5), mu0 ~ dnorm(50, pow(10,-1))

MALAVI
sigma2_0_inv ~ dgamma(5,2.5), mu0 ~ dnorm(50, pow(7.5,-1))
sigma2_0_inv ~ dgamma(5,2.5), mu0 ~ dnorm(50, pow(10,-1))

```{r}
##  ETHAN 
senate_gamma_2_2_mu_7.5.prior <- "
  nu_y ~ dunif(0,100)
  tau_y ~ dunif(0,100)
  
  mu0 ~ dnorm(50, pow(7.5,-1)) # CHANGE THIS
  
  sigma2_0 = 1/sigma2_0_inv
  sigma2_0_inv ~ dgamma(2,2) # CHANGE THIS
  "

model_file_senate_gamma_2_2_mu_7.5.prior <- generate_senate_model_text(senate_gamma_2_2_mu_7.5.prior) # CHANGE THIS
if (file.exists("senate_gamma_2_2_mu_7.5.prior.Rds")) { # CHANGE THIS
  # CHANGE THIS
  senate_gamma_2_2_mu_7.5.prior_model <- readRDS("senate_gamma_2_2_mu_7.5.prior.Rds") # CHANGE THIS
} else {
  # CHANGE THIS
  senate_gamma_2_2_mu_7.5.prior_model <- jags(data = jags_data_senate,  
                                 parameters.to.save = c("theta", "Sigma", 
                                                        "p","sigma2_y", "sigma2_0"),
                                 n.iter = 75000, n.burnin = 25000, n.thin = 50, n.chains = 2,
                                                # CHANGE THIS
                                 model.file = model_file_senate_gamma_2_2_mu_7.5.prior)
              # CHANGE THIS                                   # CHANGE THIS
  saveRDS(senate_gamma_2_2_mu_7.5.prior_model, file = "senate_gamma_2_2_mu_7.5.prior.Rds")
}
                    # CHANGE THIS
n_iter <- dim(senate_gamma_2_2_mu_7.5.prior_model$BUGSoutput$sims.list$theta[,,1])[1]

prediction_per_simulation_senate <- function(model_output, i){
  elec_sims <- model_output$BUGSoutput$sims.list$theta[i,,1]
  vote_share_by_state <- data.frame(states_senate, elec_sims)
  
  dem_states_df <- vote_share_by_state %>%
    filter(elec_sims > 50)
  dem_states <- dem_states_df$states_senate
  num_dem_states <- length(dem_states)
  return(num_dem_states)
}

final_results_senate <- c()

for (i in 1:n_iter){
                                                                      # CHANGE THIS
  final_results_senate[i] = prediction_per_simulation_senate(senate_gamma_2_2_mu_7.5.prior_model, i)
}

mean(final_results_senate >= 9)
mean(final_results_senate)
final_results_senate %>% hist()


###### NEW MODEL 
senate_gamma_2_2_mu_10.prior <- "
  nu_y ~ dunif(0,100)
  tau_y ~ dunif(0,100)
  
  mu0 ~ dnorm(50, pow(10,-1))
  
  sigma2_0 = 1/sigma2_0_inv
  sigma2_0_inv ~ dgamma(2,2)
  "

model_file_senate_gamma_2_2_mu_10.prior <- generate_senate_model_text(senate_gamma_2_2_mu_10.prior)
if (file.exists("senate_gamma_2_2_mu_10.prior.Rds")) {
  senate_gamma_2_2_mu_10.prior_model <- readRDS("senate_gamma_2_2_mu_10.prior.Rds")
} else {
  senate_gamma_2_2_mu_10.prior_model <- jags(data = jags_data_senate,  
                                 parameters.to.save = c("theta", "Sigma", 
                                                        "p","sigma2_y", "sigma2_0"),
                                 n.iter = 75000, n.burnin = 25000, n.thin = 50, n.chains = 2,
                                 model.file = model_file_senate_gamma_2_2_mu_10.prior)
  
  saveRDS(senate_gamma_2_2_mu_10.prior_model, file = "senate_gamma_2_2_mu_10.prior.Rds")
}

n_iter <- dim(senate_gamma_2_2_mu_10.prior_model$BUGSoutput$sims.list$theta[,,1])[1]

prediction_per_simulation_senate <- function(model_output, i){
  elec_sims <- model_output$BUGSoutput$sims.list$theta[i,,1]
  vote_share_by_state <- data.frame(states_senate, elec_sims)
  
  dem_states_df <- vote_share_by_state %>%
    filter(elec_sims > 50)
  dem_states <- dem_states_df$states_senate
  num_dem_states <- length(dem_states)
  return(num_dem_states)
}

final_results_senate <- c()

for (i in 1:n_iter){
  final_results_senate[i] = prediction_per_simulation_senate(senate_gamma_2_2_mu_10.prior_model, i)
}

mean(final_results_senate >= 9)
mean(final_results_senate)
final_results_senate %>% hist()
```


```{r}
# this function finds the probability that the senate flips!
#senate_gamma_4_2_prior <- readRDS("senate_gamma_2_2_prior.Rds")
n_iter <- dim(senate_gamma_4_2_prior$BUGSoutput$sims.list$theta[,,1])[1]

prediction_per_simulation_senate <- function(model_output, i){
  elec_sims <- model_output$BUGSoutput$sims.list$theta[i,,1]
  vote_share_by_state <- data.frame(states_senate, elec_sims)
  
  dem_states_df <- vote_share_by_state %>%
    filter(elec_sims > 50)
  dem_states <- dem_states_df$states_senate
  num_dem_states <- length(dem_states)
  return(num_dem_states)
}

final_results_senate <- c()

for (i in 1:n_iter){
  final_results_senate[i] = prediction_per_simulation_senate(senate_gamma_4_2_prior, i)
}

mean(final_results_senate >= 9)
mean(final_results_senate)
final_results_senate %>% hist()

#senate_gamma_4_2_prior$BUGSoutput$sims.list$theta[,14,1] %>% hist()

```

```{r}
elec_sims <- senate_gamma_5_1_prior$BUGSoutput$sims.list$theta[1,,1]
vote_share_by_state <- data.frame(states_senate, elec_sims)
```

```{r}
predictions_per_state_senate <- function(model_output, state){
  win_dem <- c()
  for (i in 1:n_iter){
    
    elec_sims <- model_output$BUGSoutput$sims.list$theta[i,,1]
    vote_share_for_state <- data.frame(states_senate, elec_sims) %>%
      filter(states_senate == state)
    vote_share_val <- vote_share_for_state$elec_sims
    if (vote_share_val > 50){
      win_dem[i] <- 1
    }
    if (vote_share_val < 50){
      win_dem[i] <- 0
    }
  }
  probability_voting_dem <- round(100*sum(win_dem)/n_iter, 3)
  return(probability_voting_dem)
}

probs_voting_dem <- c()
i = 1
for (state in states_senate){
  # print(state)
  prob_voting_dem <- predictions_per_state_senate(senate_gamma_5_1_prior, state)
  # print(prob_voting_dem)
  probs_voting_dem[i] <- prob_voting_dem
  i = i + 1
}

prob_voting_dem_table <- data.frame(states_senate, probs_voting_dem) %>%
  mutate(State = states_senate) %>%
  mutate(Probability = probs_voting_dem/100) %>%
  select(State, Probability)
kable(prob_voting_dem_table, caption = "Senate Elections: Probability of Democratic Victory in each Swing State")

```



```{r}
mean_vote_share <- senate_gamma_5_1_prior$BUGSoutput$sims.list$theta[,,1] %>% colMeans()
lower_q <- senate_gamma_5_1_prior$BUGSoutput$sims.list$theta[,,1] %>%  matrixStats::colQuantiles(probs = c(0.025))
upper_q <- senate_gamma_5_1_prior$BUGSoutput$sims.list$theta[,,1] %>%  matrixStats::colQuantiles(probs = c(0.975))

vote_share_by_state <- data.frame(states_senate, lower_q, mean_vote_share, upper_q) %>%
  mutate(State = states_senate) %>%
  mutate(`2.5%` = round(lower_q, 1)) %>%
  mutate(`97.5%` = round(upper_q, 1)) %>%
  mutate(Mean = round(mean_vote_share, 1)) %>%
  select(State, `2.5%`, Mean, `97.5%`)
kable(vote_share_by_state, caption = "Senate Elections: Predicted Democratic Vote Share in each Swing State")
```



IGNORE THIS


```{r}
states_senate <- senate$state %>% unique
y_senate <- senate$y
r_senate <- match(senate$state,states_senate)
t_senate <- senate$days_to_election + 1 
N_polls_senate <- y_senate %>% length
N_states_senate <- states_senate %>% length
N_days_senate <- t_senate %>% max
I_states_senate <- diag(N_states_senate)
jags_data_senate <- list(y=y_senate,t=t_senate,r=r_senate,
                         N_polls=N_polls_senate,N_states=N_states_senate,N_days=N_days_senate, I_states=I_states_senate)
```


```{r}
# swing states 
vote_share_2016 <- data.frame(senate$state, number = r_senate)
vote_share_2016 <- distinct(vote_share_2016) %>%
  mutate(clinton_percent = case_when(
    senate.state == "North Carolina" ~ 47/(47 + 51),
    senate.state == "Michigan" ~ 47/(47 + 48),
    senate.state == "New Hampshire" ~ 48/(48 + 47) ,
    senate.state == "Texas" ~ 43/(43 + 53) , 
    senate.state == "Minnesota" ~ 47/(47 + 45) ,
    senate.state %in% c("Georgia", "Georgia*") ~ 46/(46 + 51),
    senate.state == "Iowa" ~ 42/(42 + 52),
    senate.state == "Maine" ~ 48/(48 + 45),
    senate.state == "Arizona*" ~ 45/(48 + 45),
    senate.state == "Alabama" ~ 35/(35 + 63),
    senate.state == "Colorado" ~ 47/(47 + 44),
    senate.state == "Kansas" ~ 35/(35 + 63),
    senate.state == "South Carolina" ~ 41/(41 + 55),
    senate.state == "Kentucky" ~ 33/(33 + 63),
    senate.state == "Alabama" ~ 35/(35 + 63),
    senate.state == "Montana" ~ 36/(36 + 57),
    senate.state == "Mississippi" ~ 46.4/(53.6 + 46.4),
    senate.state == "Alaska" ~ 46.7/(53.3 + 46.7),
    senate.state == "New Mexico" ~ 54.1 / (54.1 + 30.5)
  )) %>%
  mutate(clinton_percent = 100*clinton_percent)
h <- vote_share_2016$clinton_percent
h
```

```{r}
model_mv_senate_adj_priors <- function(){
  for(k in 1:N_polls){
    y[k] ~ dnorm(p[k],1/sigma2_y[r[k]]) #note no longer binomial
    p[k] = theta[r[k],t[k]] 
  }
  for(j in 2:N_days){
    theta[1:N_states,j] ~ dmnorm(theta[1:N_states,j-1],Phi)
  }
  
  Phi ~ dwish(I_states,N_states+1) 
  Sigma = inverse(Phi)
  #hierarchical component
  for(j in 1:N_states){
    sigma2_y[j] = 1/sigma2_y_inv[j]
    sigma2_y_inv[j] ~ dgamma(nu_y,nu_y*tau_y) 
    
    #theta[j,1] ~ dnorm(h[j], pow(sigma2_0,-1))
    theta[j,1] ~ dnorm(mu0,sigma2_0)
  }
  nu_y ~ dunif(0,100)
  tau_y ~ dunif(0,100)
  
  mu0 ~ dnorm(50,pow(7.5,-1))
  
  sigma2_0 = 1/sigma2_0_inv
  sigma2_0_inv ~ dgamma(2,0.5)
}
```

```{r}
#jags_data_senate_adj_prior <- list("y","t", "r", "N_polls", "N_states", "N_days", "h", "I_states")
jags_data_senate_adj <- list(y=y_senate,t=t_senate,r=r_senate,
                             N_polls=N_polls_senate,N_states=N_states_senate,N_days=N_days_senate, I_states=I_states_senate)
jags_sims_mv_senate_adj_prior <- jags(data = jags_data_senate_adj,  model.file = model_mv_senate_adj_priors,parameters.to.save = c("theta","Sigma",
                                                                                                                                   "p","sigma2_y"),
                                      n.iter = 100000, n.burnin = 25000, n.thin = 75, 
                                      n.chains = 2
)

```


```{r}
elec_sims <- jags_sims_mv_senate_adj_prior$BUGSoutput$sims.list$theta[1,,1]
vote_share_by_state <- data.frame(states_senate, elec_sims)
```


```{r}
# this function finds the probability that the senate flips!
n_iter <- dim(jags_sims_mv_senate_adj_prior$BUGSoutput$sims.list$theta[,,1])


prediction_per_simulation_senate <- function(i){
  
  elec_sims <- jags_sims_mv_senate_adj_prior$BUGSoutput$sims.list$theta[i,,1]
  
  vote_share_by_state <- data.frame(states_senate, elec_sims)
  
  dem_states_df <- vote_share_by_state %>%
    filter(elec_sims > 50)
  dem_states <- dem_states_df$states_senate
  num_dem_states <- length(dem_states)
  return(num_dem_states)
  
}

final_results <- c()

for (i in 1:n_iter[1]){
  final_results[i] = prediction_per_simulation_senate(i)
}

mean(final_results >= 8)
(final_results) %>% hist()

# tidybayes::gather_draws(jags_sims_mv_senate_adj_prior %>% coda::as.mcmc(), theta[i, j]) %>%
#   filter(i == 1 & j == 1) %>%
#   mutate(param = ifelse(is.na(i), .variable, paste0(.variable,"[", i,", ", j, "]"))) %>%
#   ggplot(aes(x = .iteration, y = .value, color = as.factor(.chain))) +
#   geom_line(alpha = 0.6) +
#   facet_grid(param~., scales = "free_y")
```

```{r}
jags_sims_mv_senate_adj_prior
```


### Senate Visualizations 

```{r warnings=F, message=F}
h1 <- as.data.frame(senate_gamma_0.1_0.5_prior$BUGSoutput$sims.list$theta[,16,1:101]) %>%
  mutate_all(as.character())
#senate_gamma_0.1_0.5_prior$BUGSoutput$sims.list$theta[,16,1:10] # ten days before
options(dplyr.summarise.inform = FALSE)

h1 %>%
  gather("variable", "value") %>%
  mutate(variable = as.factor(variable) %>% as.numeric()) %>%
  dplyr::group_by(variable) %>% 
  #mutate(variable = as.factor(variable)) %>%
  dplyr::summarize(median_val = median(value), 
                   lower = quantile(value, probs= 0.025),
                   high = quantile(value, probs= 0.975)) %>%
  #dplyr::arrange(-row_number()) %>%
  cbind(rownames(.)) %>%
  mutate(diff = median_val - 50,
         diff_val = case_when(
           diff > 0 ~ "above",
           diff < 0 ~ "below",
           TRUE ~ "tie"
         )) %>%
  ggplot(mapping=aes(x=variable, y = diff, fill = diff_val)) +
  geom_col(position = position_stack(reverse = TRUE)) + guides(fill = FALSE) +
  labs(title = "US Senate Win Gap in North Carolina",
       subtitle = "Percent Above 50%, using 2020 Polling Data",
       caption = "The Democrat is predicted to win",
       x = "# of Days Before Election",
       y = "Percent Difference",
       fill = "") +
  scale_fill_manual( values = c("#000099", "#CC0000")) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.25, hjust=0.25, size = 7)) +
  scale_x_reverse()

z1 <- as.data.frame(senate_gamma_0.1_0.5_prior$BUGSoutput$sims.list$theta[,5,1:101]) %>%
  mutate_all(as.character())
#senate_gamma_0.1_0.5_prior$BUGSoutput$sims.list$theta[,16,1:10] # ten days before
options(dplyr.summarise.inform = FALSE)

z1 %>%
  gather("variable", "value") %>%
  mutate(variable = as.factor(variable) %>% as.numeric()) %>%
  dplyr::group_by(variable) %>% 
  #mutate(variable = as.factor(variable)) %>%
  dplyr::summarize(median_val = median(value), 
                   lower = quantile(value, probs= 0.025),
                   high = quantile(value, probs= 0.975)) %>%
  #dplyr::arrange(-row_number()) %>%
  cbind(rownames(.)) %>%
  mutate(diff = median_val - 50,
         diff_val = case_when(
           diff > 0 ~ "above",
           diff < 0 ~ "below",
           TRUE ~ "tie"
         )) %>%
  ggplot(mapping=aes(x=variable, y = diff, fill = diff_val)) +
  geom_col(position = position_stack(reverse = TRUE)) + guides(fill = FALSE) +
  labs(title = "US Senate Win Gap in Georgia Special Election",
       subtitle = "Percent Above 50%, using 2020 Polling Data",
       caption = "The Republican is predicted to win",
       x = "# of Days Before Election",
       y = "Percent Difference",
       fill = "YES") +
  scale_fill_manual(values = c("#000099", "#CC0000")) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.25, hjust=0.25, size = 7)) +
  scale_x_reverse()
```








```{r}
h1 %>%
  gather("variable", "value") %>%
  mutate(variable = as.factor(variable) %>% as.numeric()) %>%
  group_by(variable) %>%
  summarise(mean = mean(value) - 50,
            lower = quantile(value, probs=0.025) - 50,
            upper = quantile(value, probs=0.975) - 50) %>%
  mutate(party = ifelse(mean > 0, "Democrat", "Republican")) %>%
  ggplot(mapping=aes(x=variable, y = mean)) +
  geom_errorbar(aes(ymin=lower, 
                    ymax=upper,
                    color=case_when(
                      lower > 0 & upper > 0 ~ "red",
                      lower < 0 & upper < 0 ~ "blue",
                      TRUE ~ "grey"
                    ))) +
  geom_point() +
  guides(fill = FALSE) + 
  labs(title = "US Senate Win Gap in North Carolina Election",
       subtitle = "Percent Above 50%, using 2020 Polling Data",
       caption = "The Democrat is predicted to win",
       x = "# of Days Before Election",
       y = "Percent Difference",
       color=" ") +
  #scale_color_manual(values = c("#000099", "#CC0000")) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.25, hjust=0.25, size = 7)) +
  scale_x_reverse()
```


```{r}
senate_gamma_5_1_prior_v2 <- readRDS("senate_gamma_5_1_prior_v3.Rds")
n_iter <- dim(senate_gamma_5_1_prior_v2$BUGSoutput$sims.list$theta[,,1])[1]

prediction_per_simulation_senate <- function(model_output, i){
  elec_sims <- model_output$BUGSoutput$sims.list$theta[i,,1]
  vote_share_by_state <- data.frame(states_senate, elec_sims)
  
  dem_states_df <- vote_share_by_state %>%
    filter(elec_sims > 50)
  dem_states <- dem_states_df$states_senate
  num_dem_states <- length(dem_states)
  return(num_dem_states)
}

final_results_senate <- c()

for (i in 1:n_iter){
  final_results_senate[i] = prediction_per_simulation_senate(senate_gamma_5_1_prior_v2, i)
}

mean(final_results_senate >= 8)
max(final_results_senate)
sd(final_results_senate)
final_results_senate %>% hist()
senate_gamma_5_1_prior_v2$model
```

```{r}
read_csv("house_polls.csv") %>% filter(state == "North Carolina" & cycle == 2020) %>% 
  dplyr::group_by(seat_number) 
```

